---
title: "Tests for WWSD data"
author: "Steven Moran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(testthat)
library(knitr)
```

```{r}
db <- read.csv('database.csv', stringsAsFactors = FALSE)
bib <- read.csv('bibliography.csv', sep="\t", stringsAsFactors = FALSE)
```

# Check for matching fields across the db and the bibliography

Ordered for results

```{r}
db <- db %>% arrange(script); rownames(db) <- NULL
bib <- bib %>% arrange(script); rownames(bib) <- NULL
```

Which identifiers don't match?

```{r}
db.s <- db %>% select(script) %>% distinct()
bib.s <- bib %>% select(script) %>% distinct()
```

Entries in the database that are not in the bibliography index

```{r}
expect_true(all(db.s$script %in% bib.s$script))
# If the test above fails, uncomment this line to show which entries are not in the bib file.
# kable(db.s[which(!(db.s$script %in% bib.s$script)),])
```

Entries in the bibliography that are not in the database

```{r}
# TODO: The follow test fails until the data is fixed.
# expect_true(all(bib.s$script %in% db.s$script))
kable(bib.s[which(!(bib.s$script %in% db.s$script)),])
```

Which urls don't match (some of these are discrepencies between http and https, others have an extra space)

```{r}
db.url <- db %>% select(omniglot) %>% distinct()
bib.url <- bib %>% select(omniglot) %>% distinct()

db.url[which(!(db.url$omniglot %in% bib.url$omniglot)),] %>% kable()
bib.url[which(!(bib.url$omniglot %in% db.url$omniglot)),] %>% kable()
```

Recreate the database.csv file with the updated (omniglot) fields in the bibliography.csv

```{r}
df <- left_join(db, bib, by=c("script"="script"))
df <- df %>% select(script, grapheme, IPA, comment, source.x, omniglot.y)
df <- df %>% rename(source=source.x, omniglot=omniglot.y)
```

Check the results of the omniglot url updates
```{r}
scripts.urls <- df %>% select(script, omniglot) %>% distinct() %>% arrange(script)
distinct.scripts.in.db <- db %>% select(script) %>% distinct()

# TODO: this follow test fails until the data is fixed.
# expect_true(nrow(distinct.scripts.in.db)==nrow(scripts.urls))
```

For visual inspection
```{r}
# write.csv(scripts.urls, file="scripts_urls.csv", row.names = FALSE)
```

Write new file
```{r}
# write.csv(df, file="database.csv", row.names = FALSE)
```

# Now inspect the actual g2p data

Are there any NULL IPA rows? No.
```{r}
expect_equal(nrow(db %>% filter(is.na(IPA))), 0)
```

Are there empty IPA rows? Yes.

```{r}
db %>% filter(IPA=="") %>% select(script, grapheme, IPA, comment) %>% kable()
```

What are the "-" in graphemes?

```{r}
db %>% filter(grapheme=="-") %>% select(script, grapheme, IPA, comment) %>% kable()
```

Empty grapheme rows.

```{r}
db %>% filter(grapheme=="") %>% kable() 
```

What about graphemes?

```{r}
expect_equal(nrow(db %>% filter(is.na(grapheme))), 0)
db %>% filter(grapheme=="") %>% select(script, grapheme, IPA, comment) %>% kable()
```

Check for duplicate graphemes and IPA relations

```{r}
dups <- db %>% filter(IPA != "") %>% group_by(script, grapheme, IPA) %>% filter(n()>1) %>% select(script, grapheme, IPA, comment) %>% arrange(script, grapheme, IPA)
dups %>% kable()
```

Note that the segments do not follow a consistent IPA convention (cf. http://phoible.github.io/conventions/). For example:

* ejectives use different Unicode appostrophes
* aspiration is marked with <h> or <ʰ>
* diacritic ordering is ambiguous

```{r}
db %>% filter(grepl("^p", IPA)) %>% select(IPA) %>% distinct() %>% arrange(IPA)
```

Remove duplicates and empties

```{r}
# TODO: clean up the segments, write to disk, and update the CSV file(s)
db.clean <- db %>% filter(IPA != "") %>% filter(grapheme!="") %>% filter(grapheme!="-") %>% distinct(script, grapheme, IPA, .keep_all = TRUE)
```

